
# Lesson 9

```{r GPS}
getwd()
```

```{r Navigation}
setwd("/Users/kailunwang/Desktop/DAND/Case Studies")
```


```{r load data set}
library('ggplot2')
data(diamonds)
```

```{r Scatter Plot overview (qplot)}
qplot(data = diamonds, x = carat, y = price, 
      xlim = c(0, quantile(diamonds$carat, 0.99)),
      ylim = c(0, quantile(diamonds$price, 0.99))) + 
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21)

# Using a qplot to plot diamond price to against its carat weight.
# Then, trim the top percentile off both carat and price. 
```


```{r Scatter Plot overview (ggplot)}
ggplot(diamonds, aes(x = carat, y = price)) + 
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, 0.99))) + 
  scale_y_continuous(lim = c(0, quantile(diamonds$price, 0.99))) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21, alpha = 1/4) + 
  stat_smooth(method = 'lm')

```



```{r Obervations on plots}
# When we look at the plot, a few things popped out right away. 
# We can see a non-linear relationship. Maybe it's exponetial, maybe it's something else. 
# We can see that the dispersion or variance of the relationship also increases as carat size increases. 

# With just a quick look at the data, we've learned two important things about functional relationship between price and carat size. 

# We can add a linear trim line to the plot by using stat smooth function with method equals 'lm'

# We can see the linear trim line doesn't go through the center of the data at some key places. It should curve in the center of relationship and should slope up more toward the end. 

# If we tried to use this to make predictions, we might be off for some key places inside and outside of the existing data that we have displayed. 
```


```{r Conclusion about the above}
# So far, we've only consdiered bivariate relationship between price and carat weight. 

# Analyzing this data is particularly useful. Because diamonds are unique in a way that just isn't true of most manufacture products that we'used to buying. 

# You can't just plug in the model number and just look up price. 

# Hidden in the data are reflections of how a legendary marketing campaign permeated and was subsumed by our culture. 
```



```{r We shall begin to walk through our exploratory analysis}
install.packages('GGally')  # for this plot
install.packages('scales') # for variety of things
install.packages('memisc')  # for summarizing regressioin
install.packages('lattice')  # a few other things
install.packages('MASS')  # for various functions
install.packages('car')  # to recode variables
install.packages('reshape2')  # to reshape and wrangle my data
install.packages('dlyr') # to create interesting summaries and transmissions that I've done
install.packages('pastecs')

# laod ggplot graphics package and others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(pastecs)

# Sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_sample <- diamonds[sample(1:length(diamonds$price), 10000), ]

# plotting key variables against each other 
ggpairs(diamond_sample, lower = list(continuous = wrap('points', shape = I('.'))), 
        upper = list(combo = wrap('box', outlier.shape = I('.'))))
                                                                    # , axislabels = 'internal'
```

```{r Observations}
# Use histograms for qualitative, qualititave pairs
# Use scatterplots for quantitative, quantitative pairs.

# qulitative observation: a bunch of beautiful, round, and smooth glass marbles.
# quantitative observation: 5 marbles of 1.5 cm diameter each

```


```{r What correlates with prices}
# highly skewed distributions caused by monetary variables use log-transformation
```

```{r Connecting demand and distribution}
plot1 <- qplot(data = diamonds, x = price, binwidth = 100, fill = I('#099D09')) + 
  ggtitle('Price')

plot2 <- qplot(data = diamonds, x = price, binwidth = 0.01, fill = I('#F79420')) + 
  ggtitle('Log10(Price)') + 
  scale_x_log10()

library(gridExtra)
library(grid)
grid.arrange(plot1, plot2, ncol = 2)
```

## Scatter Plot Transformation
```{r}
qplot(data = diamonds, carat, price) + 
  scale_y_continuous(trans = log10_trans()) +
  ggtitle('Log10(Price) by Carat') 
```


## To make the scatter plot better, use cubic scale for volume
```{r a helper function}
cuberoot_trans = function()trans_new('cuberoot',
                                     transform = function(x) x^(1/3),
                                     inverse = function(x) x^3)
```

```{r}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3), # transform x-axis with cuberoot
                     breaks = c(0.2, 0.5, 1, 2, 3)) +
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000), # transform y-axis with log10
                     breaks = c(350, 1000, 5000, 10000, 15000)) + 
  ggtitle('Price (log10) by Cube-Root of Carat')
```

## It looks almost linear, so now we can move forward and see about modeling our data using just linear model.


## Overplotting revisited
```{r}
head(sort(table(diamonds$carat), decreasing = T)) # highest values appear first
head(sort(table(diamonds$price), decreasing = T))
```
## These numbers are too high and will cause a substantial amount of overplotting

## It looks almost linear, but surely there are other factors that influence diamond. 
## Clarity?
## Cut?

## Adjust the code below to color the points by clarity
```{r}
install.packages('RColorBrewer', dependencies = TRUE)
library(RColorBrewer)
ggplot(aes(x = carat, y = price, colour = clarity), data = diamonds) +
  geom_point(alpha = 1/2, size = 1, position = 'jitter') + 
  scale_color_brewer(type = 'div', 
                     guide = guide_legend(title = 'Clarity', reverse = TRUE, 
                                          override.aes = list(alpha = 1, size = 2))) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3), 
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000), 
                     breaks = c(350, 1000, 5000, 10000, 15000)) + 
  ggtitle('Price (log10) vs \nCube Root of Carat and Clarity')
  
```

## Clarity can explain the remaining variance in price after adding color to our plot. 



## Adjust the code below to color the points by cut
```{r}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 1/2, size = 1, position = 'jitter') + 
  scale_color_brewer(type = 'div', 
                     guide = guide_legend(title = 'Cut', 
                                          reverse = T, 
                                          override.aes = list(alpha = 1, size = 2))) +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.25, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 150000)) + 
  ggtitle('Price (log10) vs\nCube Root of Carat and Cut')
```


## Adjust the code below to color the points by color
```{r}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 1/2, size = 1, position = 'jitter') + 
  scale_color_brewer(type = 'div', 
                     guide = guide_legend(title = 'Color', 
                                          reverse = F, # Remove it, then the best will be on the top
                                          override.aes = list(alpha = 1, size = 2))) +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.25, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 150000)) + 
  ggtitle('Price (log10) vs\nCube Root of Carat and Color')
```



## Let's create linear models for price. lm(y ~ x)
```{r}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~. + carat)
m3 <- update(m2, ~. + cut)
m4 <- update(m3, ~. + color)
m5 <- update(m4, ~. + clarity)
mtable(m1, m2, m3, m4, m5)
```

## We're getting some very nice R square values. We'are accounting for almost all of the variance in price using 4 Cs(cut, color, carat, clarity). If we wanted to know whether the price for a diamond is reasonable, we might now use this model. 

## Our model is: 
## ln(Price) = 0.415 + 9.144 * carat^(1/3) - 1.093 * carat + (...coefficent...* cut + ...* color + ...* clarity) + error


## Predictions
## Example diamond from BlueNile
```{r}

thisDiamond = data.frame(carat = 1.00, cut = "Good", color = 'I', clarity = 'VS1' )

modelEstimate = predict(m5, newdata = thisDiamond, 
                        interval = 'prediction', level = 0.95)

exp(modelEstimate)

```








